\documentclass[11pt,final,twoside]{article} %setup
%set margins and load symbols and such
\usepackage[top=20mm,inner=20mm,outer=30mm,bottom=20mm,
marginparwidth=23mm,marginparsep=4mm]{geometry}
%header of document
\title{Software Development ARC Story}
\author{Grant Haataja}
\date{\today}
%set paragraph spacing
\linespread{1.5}


%make stuff happen
\begin{document}
	\maketitle

Picture an icy, bitter cold day in the middle of winter. Students shuffle from building to building on campus, some are bundled up like furry teddy bears, some are wearing only hoodies. All of them spend as little time outside as possible. From one class to the next they go, spending their time in between working on their assignments. Although fun and interesting at times, classes get old for everyone. It is hard to see the usefulness of the homework projects assigned in class. There must be something more meaningful.

From the dreams and vision of several of these unsatisfied students, ARC was born. Quickly becoming one of the largest student groups on campus, Advanced Rocketry Club gives us purpose. It gives us something to be passionate about, something we can work towards that is bigger than all of us. 

In the Software Development team for ARC, we are responsible for writing all the software code to regulate the systems of our rocket. When we started ARC last fall, most of us had no idea how to even start. Since then our team has learned how to use GitHub to effectively share our work, how to use the command-line-interface to write code on computer chips, and we have developed Open Rocket software to more accurately model our project. 

Every week we meet up in our computer lab to go over our progress and list the tasks at hand. Our next job is to write code to control the liquid-fueled engine. We have research and meetings ahead of us. We have long nights, problems, and frustration. But we have so much learning and so much excitement to experience as well, and being a part of this incredible project makes every day meaningful.


\end{document} 